\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-06-26 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

We define the \emph{positive diagonal} matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{(i,j)\in\mathcal{E}}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right),
\label{potential}	
\end{align}
wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network. The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. %The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}

@@

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular,
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.

@@

\begin{figure}[htpb]
\centering
\vspace*{-0.1in}
\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\label{fig:CompTimeSynthetic}
\end{figure}
\begin{figure}[htpb]
\centering
\vspace*{-0.1in}
\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\label{fig:RelErrSynthetic}
\end{figure}

>>>>

We define the \emph{positive diagonal} matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{(i,j)\in\mathcal{E}}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right),
\label{potential}	
\end{align}
wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology. The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. %The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}

@@

\begin{figure}[htpb]
\centering
\includegraphics[width=0.9\linewidth]{ProxSchmatic.pdf}
\caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$. The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem. While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
\vspace*{-0.1in}
\label{fig:ProxSchematic}
\end{figure}

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular,
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous. Fig. \ref{fig:ProxSchematic} illustrates the idea of approximating the joint PDF trajectory $\rho(t,\bm{x})$ through the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ computed from a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.

@@

\begin{figure}[htpb]
\centering
\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic.png}
\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:CompTimeSynthetic}
\end{figure}
\begin{figure}[htpb]
\centering
\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
\vspace*{-0.1in}
\label{fig:RelErrSynthetic}
\end{figure}
